---
title: "Course project 1"
author: "Biswajit Chowdhury"
date: "20/06/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading the data



```{r}


activityData <- read.csv("activity.csv")
head(activityData)
```



# Calculate the total number of steps taken per day

```{r}
stepPerDay<- aggregate(steps ~ date, activityData, sum)

# Make a histogram of the total number of steps taken each day
png("plot1.png", width=480, height=480)
hist(stepPerDay$steps, col="green", xlab = "Steps taken",
     main = "Total number of steps taken per day")
```


# Calculate and report the mean and median of the total number of steps taken per day

```{r}
meansteps<- mean(stepPerDay$steps, na.rm = TRUE)
mediansteps<- median(stepPerDay$steps, na.rm = TRUE)
```


#What is the average daily activity pattern?

```{r}
stepsperinterval<- aggregate(steps ~ interval, activityData, mean)

# Time series plot of the 5-minute interval and the average number of steps taken


png("plot2.png", width=480, height=480)
plot(stepsperinterval$interval, stepsperinterval$steps, type = "l", 
     xlab = "Interval", ylab = "Steps taken", 
     main = "Average Activity per day", col="red")

```

# the maximum number of steps?

```{r}
head(stepsperinterval)

maxsteps<- stepsperinterval[which.max(stepsperinterval$steps),1]
maxsteps

```


# Imputing missing values


# Calculate the total number of missing values in the data set 

```{r}
totalNa<- sum(!complete.cases(activityData))
totalNa
```

# Devise a strategy for filling in all of the missing values in the dataset. Here we fill the missing values with the mean steps taken per interval.

```{r}
new_data <- transform(activityData, steps = ifelse(is.na(activityData$steps),                                       stepsperinterval$steps[match(activityData$interval, stepsperinterval$interval)], activityData$steps))

head(new_data)

```


# histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
# Recalculate the total steps

newStepsPerDay<- aggregate(steps ~ date, new_data, sum)

# making a histogram of the the above information


png("plot1.png", width=480, height=480)
hist(newStepsPerDay$steps, xlab="Steps", 
     main="Difference between impute and non-impute data", col="red")

#Create Histogram to show difference. 


hist(stepPerDay$steps, main="Difference between impute and non-impute data",
     xlab="Steps", col="green", add=T)
legend("topleft", c("Imputed", "Non-imputed"), 
       col=c("red", "green"), lwd=10)


# Calculate new mean and median for imputed data.

newmeansteps <- mean(newStepsPerDay$steps)
newmediansteps <- median(newStepsPerDay$steps)

# Difference between new and old mean, median
diff_mean<- newmeansteps-meansteps
diff_mean

diff_med <- newmediansteps-mediansteps

diff_med

#Calculate total difference.


total_diff <- sum(newStepsPerDay$steps) - sum(stepPerDay$steps)
total_diff
```

# Are there differences in activity patterns between weekdays and weekends?

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_data$date)),
                                           
                                           weekdays), "Weekday", "Weekend"))

newStepsPerInterval <- aggregate(steps ~ interval + dow, new_data, mean)

library(lattice)
png("plot4.png", width=480, height=480)
xyplot(newStepsPerInterval$steps ~ newStepsPerInterval$interval|newStepsPerInterval$dow, 
       main="Average Steps per Day by Interval",xlab="Interval", 
       ylab="Steps",layout=c(1,2), type="l", lwd=2)
```





